---
title: "The mortality rate in Luxembourg from 1960 to 2021"
author: "Diallo Mamadou Korka"
date: "2022-11-25"
output:
  html_document:
    css: "styles.css"
  pdf_document: default
  
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r ,fig.width=8, fig.height=8, comment = NA, echo=FALSE, message = FALSE, warning = FALSE}
library(readr)
library(tidyselect)
library(tidyverse)
library(broom)
library(vroom)
library(gganimate)
library(maps)
library(cowplot)
library(tinytex)
library(plotly)
library(viridis)
library(gifski)


```



<div class="card">
<div class="card-header">
Data Set
</div>
<blockquote>
The professor suggested that it would be more efficient to download the datasets directly using the provided links on the website. However, the source of the datasets was set up in a way that did not allow for copying the links. As a result, I had to download the datasets directly from the website. The datasets used in this study include: 'What causes contribute to mortality rates in Luxembourg', 'Non communication in Luxembourg', and 'Population by age group'. The links for these datasets are as follows: <a href="https://platform.who.int/mortality/countries/country-details/MDB/luxembourg">What causes contribute to mortality rates in Luxembourg, from the website "platform.who" </a>, <a href="https://platform.who.int/mortality/themes/theme-details/MDB/noncommunicable-diseases">Non communication in Luxembourg</a>, and <a href="https://ourworldindata.org/age-structure">Population by age group, source "ourworldindata.org"</a>.
</blockquote>


```{r ,fig.width=8, fig.height=8, comment = NA, echo=FALSE, message = FALSE, warning = FALSE}

#Mortality in Luxembourg
Mortality<-read_csv("data/Mortality_Database_Luxembourg.csv",
                    skip = 9,
    
        
      col_names  = c( "Indicator_Code",
                     "Indicator_Name",
                     "Year",
                     "Sex",
                     "Age_group_code",
                     "Age_Group",
                     "Number",
                     "Percentage_of_cause_specific_deaths_out_of_total_deaths",
                     "Age_standardized_death_rate_per_100000_standard_population",
                     "Death_rate_per_100000_population",
                     "Dumm"),
         show_col_types = TRUE)


#Mortality causes by noncauminicate desease in the in luxembourg from 1967
Noncauminicate_in_luxembourg<-read_csv("data/Luxembourg/WHOMortalityDatabase_Trends_years_sex_a_country_area_age-Noncommunicable diseases_1st January 2023 17_38.csv",
                         skip = 7,
                         col_names  = c( "Region_Code",
                     "Region_Name",
                     "Country_Code",
                     "region",
                     "Year",
                     "Sex",
                     "Age_group_code",
                     "Age_Group",
                     "Number",
                     "Percentage_of_cause_specific_deaths_out_of_total_deaths",
                     "Age_standardized_death_rate_per_100_000_standard_population",
                     "Death_rate_per_100_000_population",
                     ""),
         show_col_types = TRUE)
#Noncauminicate_in_luxembourg

#Mortality causes by noncauminicate desease in the world
Noncauminicate<-read_csv("data/WHOMortalityDatabase_Trends_years_many_countries_by_age_sex-Noncommunicable diseases_7th January 2023 03_11.csv",
                         skip = 7,
                         col_names  = c( "Region_Code",
                     "Region_Name",
                     "Country_Code",
                     "region",
                     "Year",
                     "Sex",
                     "Age_group_code",
                     "Age_Group",
                     "Number",
                     "Percentage_of_cause_specific_deaths_out_of_total_deaths",
                     "Age_standardized_death_rate_per_100_000_standard_population",
                     "Death_rate_per_100_000_population",
                     ""),
         show_col_types = TRUE)
#Noncauminicate

#Population by group age Europe
Population<-read_csv("data/age_dependency_ratio_of_working_age_population.csv",
                         skip = 7,
                         col_names  = c( "region",
                     "Code",
                     "Year",
                     "estimates_age",
                     ""),
         show_col_types = TRUE)
#Population

```

</div> 
<br/><br/>




<div class="card">
<div class="card-header">
Examining Temporal Trends and Gender Disparities in Mortality Rates, 1960-2021.
</div>
  

<div class="card-body">
The objective of this study is to investigate the temporal dynamics of mortality rates from 1960 to 2021, encompassing individuals of all age groups. A key focus of the analysis will be to assess potential disparities in mortality rates between males and females.
        

```{r , comment = NA, echo=FALSE, message = FALSE, warning = FALSE}


Mortality %>% filter(Age_group_code == "Age_all", Indicator_Name=="All Causes" ) %>% 

ggplot(aes(x=Year, y=Death_rate_per_100000_population, group=Sex))+
  geom_line(aes(color=Sex))+
  geom_point(aes(color=Sex))+
  xlab("Year") + ylab("Death rate per 100,000 population") +
  ggtitle("Mortality rate over years in Luxembourg")+
  transition_reveal(Year)

```
</div>

<div class="card-footer">
<blockquote>
We notice that in Luxembourg, the mortality rate was very high in 1960.because in Luxembourg there have been several epidemics that could affect lives.However, from 1960 to 2020, this rate has decreased. This makes sense, because Luxembourg has adopted more modern treatment methods.
 </blockquote>
</div>

</div>

<br/><br/>

<div class="card">
<div class="card-header">
We shall now analyze some major causes of mortality in luxembourg and how these causes have affected different age groups.
</div>
  
<div class="card-body">


```{r, comment = NA, echo=FALSE, message = FALSE, warning = FALSE}
Mortality |> count(Indicator_Name, sort = TRUE)

ggplot(data = Mortality |> filter(Number>1000 , Indicator_Name != "All Causes")) + 
  geom_col(aes(y = Number, x = Year, fill=Indicator_Name)) +
  facet_wrap(~Indicator_Name)

```

</div>

<div class="card-footer">
<blockquote>
We have identified some of the major causes of mortality in Luxembourg over the years. According to our data, the number of deaths due to non-communicable diseases is quite high.

Deaths from cardiovascular diseases and malignant neoplasm have been steadily decreasing, which may be due to improvements in sanitation infrastructure.

malignant neoplasm deaths appeared to be low in the years 2004, but then saw an increase around 2005 until 2021. This could be due to outbreaks of malignant neoplasm or the aging of the Luxembourg population.

It is worth noting that the population of Luxembourg has been aging over this time period.

</blockquote>

  </div>
</div>
<br/><br/>


<div class="card">
<div class="card-header">
Given that non-communicable diseases are the leading cause of death,it's important to visualize the impact of these conditions on mortality by age group. This will allow us to identify which age brackets are most affected.
</div> 
<div class="card-body">

```{r, comment = NA, echo=FALSE, message = FALSE, warning = FALSE}
#summurase
#(?<=\\[).*(?=\\])

#Mortality

Mortality %>% filter(Indicator_Name =="Noncommunicable diseases", Age_group_code != "Age_all") %>%
  mutate(Age_by_group_New = str_extract(Age_Group,"(\\d)+") %>% as.numeric(),
         Age_by_group_New = if_else(Age_group_code == "Age_all", 74, Age_by_group_New),
         Age_Group = fct_reorder(Age_Group, Age_group_code)) %>%
  
  ggplot(aes(x = Age_Group, y=Death_rate_per_100000_population, fill=Sex))+
  geom_col(size=2, alpha=0.5)+
  #geom_line(size=1)+
  theme_minimal()+ 
  
  xlab("Age Group") + ylab("Death rate per 100,000 population") +
  ggtitle("Cumulative Mortality from Non-Communicable Diseases in Luxembourg by Year.")+
  
  theme(axis.text.x = element_text(angle = 90))
```
</div>

<div class="card-footer">
 <blockquote>
Here we plot the number of deaths from non-communication diseases for different age groups. From the graph, we can understand that the number of deaths due to non_communication issues is maximum for the age groups from 0 and also age groups between 54 and 74 years old. We also note that the number of deaths due to non-communication of illnesses is down slightly for the young population. The top no.

child deaths may be due to the fact that children of this age are not strong enough to resist this disease. It is noted that the number of deaths is also relatively higher among the elderly than among middle-aged adults. This may be the case with older people who do not have strong immunity against non-communication diseases.
 </blockquote>
</div>
</div>

<br/><br/>


<div class="card">
<div class="card-header">
We have determined that this disease disproportionately affects older individuals. As such, it is important to visualize how this disease has contributed to mortality trends from 1960 to 2021.
</div>
<div class="card-body">
```{r, comment = NA, echo=FALSE, message = FALSE, warning = FALSE}
base <- Noncauminicate_in_luxembourg %>% filter(Sex == "All") %>% 
  plot_ly(x = ~Number, y = ~Age_Group, size = ~3, 
          text = ~Age_Group, hoverinfo = "text") %>%
  layout(xaxis = list(type = "log"))

base %>%
  add_markers(color = ~Age_Group, frame = ~Year, ids = ~Age_Group) %>%
  animation_opts(1000, easing = "elastic", redraw = FALSE) %>%
  animation_button(
    x = 1, xanchor = "right", y = 0, yanchor = "bottom"
  ) %>%
  animation_slider(
    currentvalue = list(prefix = "YEAR ", font = list(color="red"))
  )
```
</div>
<div class="card-footer">
 <blockquote>
Our analysis has shown that the 70-year age group has consistently had the highest number of mortality cases from 1967 to 1980.
 </blockquote>
</div>

</div>

<br/><br/>


<div class="card">
<div class="card-header">
To fully understand the impact of non-communicable diseases on the 70+ population in Luxembourg, it is necessary to compare the mortality rates in this age group to those of similar populations in other countries
</div>
  
<div class="card-body" id="img_map">
```{r ,fig.width=8, fig.height=8, comment = NA, echo=FALSE, message = FALSE, warning = FALSE}

mapdata <- map_data("world")
newDataForNonCauminicate<-Noncauminicate %>% 
                                            select(region,Year, Number,Death_rate_per_100_000_population, Age_Group,Age_group_code)  %>% 
                                            filter(Age_group_code == "Age70_74", (Year >= 1967 & Year <= 1980)) 
 
mapdata_join<-left_join(mapdata,newDataForNonCauminicate, by="region")
map1<- ggplot(mapdata_join, aes(x = long, y = lat,fill=Number, group=group))+
  geom_polygon(color="black")+ 
  theme(panel.background = element_rect(fill = "#101045")) +
  coord_equal()

#map1

map2<-map1 + scale_fill_gradient(name="%Number of dies", low = "Yellow", high="red")+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        rect = element_blank()
        )

map2

```
</div>



<div class="card-footer">
 <blockquote>
Our analysis has shown that the 70-year age group has consistently had the highest number of mortality cases from 1967 to 1980 in Europe.
 </blockquote>
</div>

</div>
<br/><br/>


<div class="card">
<div class="card-header">
Checking the age categories of the population in Europe from 1967 to 1980 could be important for understanding the prevalence of non-communicable diseases (NCDs) in this population during that time period.

This is because NCDs, such as heart disease, stroke, and cancer, are often more common in older age groups. By examining the age distribution of the population, We can determine whether the observed prevalence of NCDs in the population is due to the higher proportion of older individuals in the population, or whether other factors are also contributing to the observed prevalence of NCDs.

</div>
  
<div class="card-body">

```{r ,fig.width=8, fig.height=8, comment = NA, echo=FALSE, message = FALSE, warning = FALSE}
# Define a vector of European countries
european_countries <- c("Albania", "Andorra", "Armenia", "Austria", "Azerbaijan", "Belarus", "Belgium", "Bosnia and Herzegovina", "Bulgaria", "Croatia", "Cyprus", "Czech Republic", "Denmark", "Estonia", "Finland", "France", "Georgia", "Germany", "Greece", "Hungary", "Iceland", "Ireland", "Italy", "Kazakhstan", "Kosovo", "Latvia", "Liechtenstein", "Lithuania", "Luxembourg", "Macedonia", "Malta", "Moldova", "Monaco", "Montenegro", "Netherlands", "Norway", "Poland", "Portugal", "Romania", "Russia", "San Marino", "Serbia", "Slovakia", "Slovenia", "Spain", "Sweden", "Switzerland", "Turkey", "Ukraine", "United Kingdom", "Vatican City")

gap <- Population %>% filter((Year >= 1967 & Year <= 1980), region %in% european_countries) %>% 
  group_by(Year) %>%
  # The * 1 makes it possible to have non-integer ranks while sliding
  mutate(rank = min_rank(-estimates_age) * 1) %>%
  ungroup()

p <- ggplot(gap, aes(rank, group = region, 
                     fill = as.factor(region), color = as.factor(region))) +
 


  geom_tile(aes(y = estimates_age/2,
                height = estimates_age,
                width = 0.9), alpha = 0.8, color = NA) +



  # text in x-axis (requires clip = "off" in coord_*)
  # paste(country, " ")  is a hack to make pretty spacing, since hjust > 1 
  #   leads to weird artifacts in text spacing.
  geom_text(aes(y = 0, label = paste(region, " ")), vjust = 0.2, hjust = 1) +

  coord_flip(clip = "off", expand = FALSE) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_reverse() +
  guides(color = FALSE, fill = FALSE) +

  

  labs(title='{closest_state}', x = "", y = "Age") +
  
  theme(plot.title = element_text(hjust = 0, size = 22),
        axis.ticks.y = element_blank(),  # These relate to the axes post-flip
        axis.text.y  = element_blank(),  # These relate to the axes post-flip
        plot.margin = margin(1,1,1,4, "cm")) +
  

  transition_states(Year, transition_length = 4, state_length = 1) +
  
  ease_aes('cubic-in-out')

animate(p, fps = 25, duration = 20, width = 800, height = 600)

```
</div>
<div class="card-footer">
 <blockquote>
 
 Based on this visualization, it is likely that you will observe that the median age of the population in some countries was higher than in Luxembourg during the 1967-1980 time period. This suggests that mortality due to non-communicable diseases was not solely caused by the aging of the population in these countries, but may also have been influenced by other factors such as genetics, lifestyle, and environmental exposures. From this information, it is possible to conclude that mortality due to non-communicable diseases was not only caused by the aging of the population in these countries during this time period.
 
 
 </blockquote>
</div>

</div>
<br/><br/>



<div class="card">
<div class="card-header">
Thank you
</div>
<blockquote>

<p>I am deeply grateful to <b>Professor Roland Krause and his assistants</b> for their invaluable guidance throughout my data visualization project.</p>
<p>I utilized various websites as resources for obtaining the datasets used in my project.</p>
<ul>
<li><a href="https://platform.who.int/mortality/countries/country-details/MDB/luxembourg">platform.who</a> </li>
<li><a href="https://platform.who.int/mortality/themes/theme-details/MDB/noncommunicable-diseases">Non communication in Luxembourg</a> </li>
<li><a href="https://ourworldindata.org/age-structure">ourworldindata.org</a> </li>
</ul>



<br/>

<p>I would like to acknowledge the use of various websites as helpful resources throughout my data visualization project.
<ul>
<li>course material</li>
<li><a href="https://www.r-project.org/about.html">r-project</a> </li>
<li><a href="https://stackoverflow.com/">stack overflow</a> </li>
<li><a href="https://developer.mozilla.org/en-US/docs/Web/CSS">css</a> </li>
</ul>
</p>

</blockquote>

</div>
<br/>
<br/>
<br/>


