---
title: "mortality_indicators_luxembourg"
author: "Diallo Mamadou Korka"
date: "2022-11-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#library

```{r}
library(readr)
library(tidyselect)
library(tidyverse)
library(broom)
library(vroom)
library(gganimate)

```

#Link to Download the dataset https://platform.who.int/mortality/countries/country-details/MDB/luxembourg
#Link to Download the dataset world https://platform.who.int/mortality/themes/theme-details/MDB/noncommunicable-diseases
#https://ourworldindata.org/age-structure
#Read the dataset and Tidy

```{r}

#Mortality in Luxembourg
Mortality<-read_csv("data/Mortality_Database_Luxembourg.csv",
                    skip = 9,
    
        
      col_names  = c( "Indicator_Code",
                     "Indicator_Name",
                     "Year",
                     "Sex",
                     "Age_group_code",
                     "Age_Group",
                     "Number",
                     "Percentage_of_cause_specific_deaths_out_of_total_deaths",
                     "Age_standardized_death_rate_per_100000_standard_population",
                     "Death_rate_per_100000_population",
                     "Dumm"),
         show_col_types = TRUE)

Mortality

#Mortality causes by noncauminicate desease
Noncauminicate<-read_csv("data/WHOMortalityDatabase_Map_Noncommunicable diseases_14th_December_2022_17_17.csv",
                         skip = 7,
                         col_names  = c( "Region_Code",
                     "Region_Name",
                     "Country_Code",
                     "region",
                     "Year",
                     "Sex",
                     "Age_group_code",
                     "Age_Group",
                     "Number",
                     "Percentage_of_cause_specific_deaths_out_of_total_deaths",
                     "Age_standardized_death_rate_per_100_000_standard_population",
                     "Death_rate_per_100_000_population",
                     ""),
         show_col_types = TRUE)
Noncauminicate

#Population by group age Europe
Population<-read_csv("data/age_dependency_ratio_of_working_age_population.csv",
                         skip = 7,
                         col_names  = c( "region",
                     "Code",
                     "Year",
                     "Total dependency ratio - Sex: all - Age: none - Variant: estimates",
                     ""),
         show_col_types = TRUE)
Population

```

#We want to see how the mortality rate varies over the course of 1960 to 2021. 
```{r}


Mortality %>% filter(Age_group_code == "Age_all", Indicator_Name=="All Causes" ) %>% 

ggplot(aes(x=Year, y=Number, group=Sex))+
  geom_line(aes(color=Sex))+
  geom_point(aes(color=Sex))+
  xlab("Year") + ylab("Number") +
  ggtitle("Mortality rate over years in Luxembourg")


```

#We shall now analyze a few major causes of mortality in luxembourg and how these causes have affected different parts of the sex\

```{r}
Mortality |> count(Indicator_Name, sort = TRUE)

ggplot(data = Mortality |> filter(Number>1000)) + 
  geom_col(aes(y = Number, x = Year)) +
  facet_wrap(~Indicator_Name)

```

#Number of deaf by noncommunicatif desease by cat age
```{r}
#summurase
#(?<=\\[).*(?=\\])

#Mortality

Mortality %>% filter(Indicator_Name =="Noncommunicable diseases") %>%
  mutate(Age_by_group_New = str_extract(Age_Group,"(\\d)+") %>% as.numeric(),
         Age_by_group_New = if_else(Age_group_code == "Age_all", 74, Age_by_group_New),
         Age_Group = fct_reorder(Age_Group, Age_group_code)) %>%
  
  ggplot(aes(x = Age_Group, y=Number, color=Sex))+
  geom_point(size=2, alpha=0.5)+
  #geom_line(size=1)+
  theme_minimal()+
  labs(title="Yearly Culmulative mortality by NonComunication deseses in Luxembourg")
  #transition_reveal(Year)
 


```

#Find the Year , where we have most mortality from Age 54 to 74 in luxembourg
```{r}
Mortality %>%filter(Age_group_code %in% c("Age55_59","Age60_64","Age65_69","Age70_74"), Indicator_Name=="Noncommunicable diseases") %>% 

  ggplot(aes(x = Year, y=Number, fill=Age_group_code))+
    geom_col()
    #geom_point()
  


```

#Comparing the mortality cause by noncomminicate desase in Luxembourg to other country from 54 to 74

```{r}


mapdata <- map_data("world")


#mapdata
#Noncauminicate %>% filter(Age_group_code == "Age75_79") %>% 
#  select (region, Number) %>% 
  
#mapdata<- left_join(mapdata,Noncauminicate, by="region")


```

#Life experience in 2000 comparing the rest of the word in 2000(Map) with noncommunication desases

```{r}


```

#study cas of the mortality by noncommunicate deseas(Regression linear)

```{r}
ggplot(data = Mortality |> filter(Indicator_Name =="Noncommunicable diseases")) + 
  geom_point(aes(y =Number , x = Age_Group, fill=Sex)) #+
  #geom_bar()
  
  #facet_wrap(~Sex)

```

#Prediction for 2023

```{r}

```
